---
title: "MSBA 6420 - Predictive Analytics Project"
author: "Lavanya Basava Raju, Ananya Mishra, Xiurong Lin, Ryan Borowicz"
date: "December 5, 2016"
output: html_document
---

## SETUP
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require("knitr")
opts_knit$set(root.dir = "C:/Users/Lavanya/Google Drive/Group Project/Case1-MSRA/Training/")
library(dplyr) 
library(tidyr)
library(data.table)
library(ggplot2)
library(magrittr)
library(lubridate)
library(fastcluster)
library(reshape2)

folder="C:/Users/Lavanya/Google Drive/Group Project/Case1-MSRA/Training/"
setwd(folder)
```

## BUSINESS UNDERSTANDING:
Overall Goal:
* Part 1. Predict who will get pneumonia & mrsa
* Part 2. Finding the people who are likely to contract nosocomial pneumonia, and figuring out how much it costs to treat them for this infection. We can't give you the rule to create target labels in Part 2, since a) there isn't a perfect one, and b) creating a good one is part of the contest.

MRSA & Pneumonia:
* MRSA is represented by ICD9 code V09.0. It can be identified either in the hospital file using one of the ICD9 fields, or in the patient conditions file, which lists all conditions for which a patient was diagnosed for that year.
* There is a cost of preventive care that you can determine from the prescription database. It generally requires the antibiotic, Vancomycin or the antibiotic, Zyvox (linezolid). Then there is the added cost of MRSA, which is not uniform; rather, it depends upon the patient condition and the patient procedures performed. You will need to estimate this added cost from the data. Both emergency and elective patients have some risk of MRSA. The hospital will want to use the preventive treatment if it costs less than the added cost of MRSA.
* There are 971 patients in the conditions data set with at least one ICD9CODX value between 480 and 486. In the hospital data set, there are 278 unique DUPERSIDs with at least one of Icd1x, Icd2x, Icd3x, and Icd4x between 480 and 486. (These 278 patients all appear to be in the conditions data base as well, with a code of pneumonia for at least one record.)

Other Information:
* File RX0304 is CSV with quoted fields if a comma is part of the field - make sure we covered this

* 9 - Not ascertained - not NA
* 8 - Unknown - NA
* 7 - Refused to Answer - not NA
* 1 - Inapplicable - convert to NA

## CONDITIONS DATASET
* not one to one with hospital data - conditions without associated record in hospital table - ex. 20003019 w/6 condition rows - only 1 hospital stay for 1 of 4 conditions - contains many more patients than hospitals dataset
* not one to one with rx dataset - rx could have medications for 1 of many conditions

```{r CONDITIONS}
conditions <- read.csv("conditions0304.csv", header = TRUE,na.strings = c("-8","-1"))

#Change -9 ICD values to ND - Different from NA - these are conditions that can't be accurately diagnosed at hospital so they are not given an ICD code?
conditions$ICD9CODX <- as.character(conditions$ICD9CODX)
conditions$ICD9CODX[conditions$ICD9CODX == "-9"] <- 'ND'
conditions$ICD9CODX <- as.factor(conditions$ICD9CODX)

#convert year to factor
conditions$year <- as.factor(conditions$year)

#check for missing values & duplicates
str(conditions)
summary(conditions)  # missing values in icd codes

#remove missing values
conditions %<>% filter(!is.na(ICD9CODX)) 
summary(conditions)

colnames(conditions)
# "DUID"     "DUPERSID" "ICD9CODX" "year"
nrow(conditions)
conditions_unique_records <- conditions %>% group_by(DUID, DUPERSID, ICD9CODX, year) %>% summarize(count=n())
nrow(conditions_unique_records)   # table has duplicates
conditions_2 <- conditions_unique_records # conditions datasets without duplicates
conditions_2_unique_records <- conditions_2 %>% group_by(DUPERSID, ICD9CODX, year) %>% summarize(count=n())
nrow(conditions_2_unique_records) # PK - DUPERSID, ICD9CODX, year
rm(conditions_unique_records)
rm(conditions_2_unique_records)

#Remove DUID - info already captured in DUPERSID
myvars <- c("DUPERSID", "ICD9CODX", "year")
conditions_3 <- conditions_2[myvars]

rm(conditions,conditions_2)
summary(conditions_3)
```

## DEMOGRAPHICS DATASET
```{r DEMOGRAPHICS}
demographics <- read.csv("demographics0304.csv", header = TRUE,,na.strings = c("-8","-1"))

str(demographics)
#DUID & PID will be removed - covered by DUPERSID
#convert DOBMM and DOBYY to factor
demographics$DOBMM <- as.factor(demographics$DOBMM)
demographics$DOBYY <- as.factor(demographics$DOBYY)

#convert sex to factor
demographics$SEX[demographics$SEX == 1] <- 'Male'
demographics$SEX[demographics$SEX == 2] <- 'Female'
demographics$SEX <- as.factor(demographics$SEX)

#RACEX
demographics$RACEX[demographics$RACEX == 1] <- 'White'
demographics$RACEX[demographics$RACEX == 2] <- 'Black'
demographics$RACEX[demographics$RACEX == 3] <- 'Native'
demographics$RACEX[demographics$RACEX == 4] <- 'Asian'
demographics$RACEX[demographics$RACEX == 5] <- 'Hawaiian'
demographics$RACEX[demographics$RACEX == 6] <- 'Multiple'

#use information from other race columns to classify - RACEAX (asian), RACEBX (black), RACEWX (white), RACETHNX, HISPANX (hispanic), HISPCAT (didn't break out hispanic group by category (i.e. cuban, mexican, dominican)
#re-classify hispanics into own category - initially marked on survey as white category
demographics$RACEX[demographics$RACEX == 'White' & demographics$HISPANX == 1] <- 'Hispanic'
demographics$RACEX <- as.factor(demographics$RACEX)

#education year - remove or group into college, high-school, non high-school - values don't seem to be very trustworthy
#changing to 0 probably isn't accurate - more likely NA
demographics$EDUCYEAR[is.na(demographics$EDUCYEAR)] <- 0
demographics$EDUCYEAR[demographics$EDUCYEAR == -9] <- 0
demographics$EDUCYEAR[demographics$EDUCYEAR == -7] <- 0
demographics$EDUCYEAR <- as.factor(demographics$EDUCYEAR)

#convert year to factor
demographics$Year <- as.factor(demographics$Year)

#marry
demographics$marry[demographics$marry == 1] <- 'Married'
demographics$marry[demographics$marry == 2] <- 'Widowed'
demographics$marry[demographics$marry == 3] <- 'Divorced'
demographics$marry[demographics$marry == 4] <- 'Separated'
demographics$marry[demographics$marry == 5] <- 'Never Married'
demographics$marry[demographics$marry == 6] <- 'Under16'

#convert -7, -9, NA values to Never Married
demographics$marry[demographics$marry == -7] <- 'Never Married'
demographics$marry[demographics$marry == -9] <- 'Never Married'
demographics$marry[is.na(demographics$marry)] <- 'Never Married'

demographics$marry <- as.factor(demographics$marry)
#income - alot of 0's - have to be careful using this variable - probably not that accurate
#convert negative income values to positive - assuming a misplaced negative sign
demographics$income <- abs(demographics$income)

#poverty - doesn't correlate with income - not sure what this is representing - remove
demographics$poverty[demographics$poverty == 1] <- 'Poor'
demographics$poverty[demographics$poverty == 2] <- 'Nearpoor'
demographics$poverty[demographics$poverty == 3] <- 'Lowincome'
demographics$poverty[demographics$poverty == 4] <- 'Midincome'
demographics$poverty[demographics$poverty == 5] <- 'Highincome'
demographics$poverty <- as.factor(demographics$poverty)

str(demographics)
summary(demographics)
colnames(demographics)
# "DUID"     "PID"      "DUPERSID" "DOBMM"    "DOBYY"    "SEX"      "RACEX"    "RACEAX"   "RACEBX"   "RACEWX"  
# "RACETHNX" "HISPANX"  "HISPCAT"  "EDUCYEAR" "Year"     "marry"    "income"   "poverty" 

#keep only important variables
myvars <- c("DUPERSID", "DOBMM", "DOBYY", "SEX", "RACEX", "EDUCYEAR", "Year", "marry", "income")
demographics_2 <- demographics[myvars]

nrow(demographics_2)
demographics_unique_records <- demographics_2 %>% group_by(DUPERSID, Year) %>% summarize(count=n())
nrow(demographics_unique_records)   # PK - DUPERSID, Year
rm(demographics_unique_records)

colnames(demographics_2) <- c(
"DUPERSID", "DOBMM", "DOBYY", "SEX", "RACEX", "EDUCYEAR", "year", "marry", "income")

rm(demographics)
summary(demographics_2)
```

## HOSPITAL DATASET

* 9 - Not ascertained - not NA
* 8 - Unknown - NA
* 7 - Refused to Answer - not NA
* 1 - Inapplicable - convert to NA

DUPERSID	PERSON ID (DUID + PID)


Icd1x	Primary diagnosis
Icd2x	Secondary diagnosis
Icd3x	Secondary diagnosis
Icd4x	Secondary diagnosis

ANYOPER	Any operations or surgeries performed
Pro1x	Primary procedure
Pro2x	Secondary procedure

ERHEVIDX	Event ID for corresponding emergency room visit
EMERROOM	Did stay begin with emergency room visit?
RSNINHOS	Reason entered hospital
SPECCOND	Hospital stay related to condition
NUMNIGHT	Number of nights stayed at provider
NUMNIGHX	# of nights in hospital-edited/imputed
DSCHPMED	Medicines prescribed at discharge

VAPLACE	VA Facility flag

Mdpayment	Physician reimbursement
Mdtotal	Total physician cost
Fcpayment	Facility payment
Fctotal	Total facility cost
Totalcharge	Total charges
Totalexpenditure	Total reimbursement 

```{r HOSPITAL}
hospital <- read.csv("hospital0304.csv", header = TRUE, sep=",", na.strings = '.')
str(hospital)
summary(hospital)


#ERHEVIDX - what does -1 mean vs blank?
#emergency room values not consistent - event id's where emergency room = 2(no)
#if there is an emergency event id but emerroom=2(no), count is as false alarm - not emergency room stay - imputed nights should be changed to 0 for these records
#bad data - remove

hospital$EMERROOM[hospital$EMERROOM == 1] <- 1
hospital$EMERROOM[hospital$EMERROOM == 2] <- 0
hospital$EMERROOM[is.na(hospital$EMERROOM)] <- 0
hospital$EMERROOM <- ifelse(hospital$EMERROOM < 0,0,hospital$EMERROOM)
hospital$EMERROOM <- ifelse(hospital$EMERROOM > 1,0,hospital$EMERROOM)

#icd values - convert -9 to nd
hospital$icd1x <- as.character(hospital$icd1x)
hospital$icd1x[hospital$icd1x == "-9"] <- 'ND'
hospital$icd1x[hospital$icd1x == "-1"] <- 'NA'
hospital$icd1x <- as.factor(hospital$icd1x)

hospital$icd2x <- as.character(hospital$icd2x)
hospital$icd2x[hospital$icd2x == "-9"] <- 'ND'
hospital$icd2x[hospital$icd2x == "-1"] <- 'NA'
hospital$icd2x <- as.factor(hospital$icd2x)

hospital$icd3x <- as.character(hospital$icd3x)
hospital$icd3x[hospital$icd3x == "-9"] <- 'ND'
hospital$icd3x[hospital$icd3x == "-1"] <- 'NA'
hospital$icd3x <- as.factor(hospital$icd3x)

hospital$icd4x <- as.character(hospital$icd4x)
hospital$icd4x[hospital$icd4x == "-9"] <- 'ND'
hospital$icd4x[hospital$icd4x == "-1"] <- 'NA'
hospital$icd4x <- as.factor(hospital$icd4x)

#procedure values
hospital$pro1x <- as.character(hospital$pro1x)
hospital$pro1x[hospital$pro1x == "-9"] <- 'ND'
hospital$pro1x[hospital$pro1x == "-1"] <- 'NA'
hospital$pro1x <- as.factor(hospital$pro1x)

hospital$pro2x <- as.character(hospital$pro2x)
hospital$pro2x[hospital$pro2x == "-9"] <- 'ND'
hospital$pro2x[hospital$pro2x == "-1"] <- 'NA'
hospital$pro2x <- as.factor(hospital$pro2x)

#nights - need to decide how to handle this
#use NUMNIGHT if not negative, use NUMNIGHX (imputed value) if negative
#what does -1,-7,-8,-9 mean? Different than blank?
hospital$NUMNIGHX <- as.integer(hospital$NUMNIGHX)
hospital$NUMNIGHT <- as.integer(hospital$NUMNIGHT)
hospital$num_nights = ifelse(hospital$NUMNIGHT < 0,hospital$NUMNIGHX,hospital$NUMNIGHT)

#Hospital stay related to condition?
#what does -1,-7,-8,-9 mean? Different than blank?
#change na's to no?
hospital$SPECCOND[hospital$SPECCOND == 1] <- 1
hospital$SPECCOND[hospital$SPECCOND == 2] <- 0
hospital$SPECCOND[hospital$SPECCOND == -1] <- 0
hospital$SPECCOND[hospital$SPECCOND == -7] <- 0
hospital$SPECCOND[hospital$SPECCOND == -8] <- 0
hospital$SPECCOND[hospital$SPECCOND == -9] <- 0
hospital$SPECCOND[is.na(hospital$SPECCOND)] <- 0
hospital$SPECCOND <- ifelse(hospital$SPECCOND < 0,0,hospital$SPECCOND)
hospital$SPECCOND <- ifelse(hospital$SPECCOND > 1,0,hospital$SPECCOND)
hospital$SPECCOND <- as.numeric(hospital$SPECCOND)

#Reason entered hospital
#what does -1,-7,-8,-9 mean? Different than blank?
#change na's to other?
hospital$RSNINHOS[hospital$RSNINHOS == 1] <- 'Surgery'
hospital$RSNINHOS[hospital$RSNINHOS == 2] <- 'Treatment'
hospital$RSNINHOS[hospital$RSNINHOS == 3] <- 'Diagnostic'
hospital$RSNINHOS[hospital$RSNINHOS == 4] <- 'Babybirth'
hospital$RSNINHOS[hospital$RSNINHOS == 5] <- 'Babyprebirth'
hospital$RSNINHOS[hospital$RSNINHOS == 91] <- 'Other'
hospital$RSNINHOS[is.na(hospital$RSNINHOS)] <- 'Other'
hospital$RSNINHOS[hospital$RSNINHOS == -1] <- 'Other'
hospital$RSNINHOS[hospital$RSNINHOS == -7] <- 'Other'
hospital$RSNINHOS[hospital$RSNINHOS == -8] <- 'Other'
hospital$RSNINHOS[hospital$RSNINHOS == -9] <- 'Other'
hospital$RSNINHOS <- as.factor(hospital$RSNINHOS)

#Any operations or surgeries performed
#what does -1,-7,-8,-9 mean? Different than blank?
#change na's to no?
hospital$ANYOPER[hospital$ANYOPER == 1] <- 1
hospital$ANYOPER[hospital$ANYOPER == 2] <- 0
hospital$ANYOPER[is.na(hospital$ANYOPER)] <- 0
hospital$ANYOPER <- ifelse(hospital$ANYOPER < 0,0,hospital$ANYOPER)
hospital$ANYOPER <- ifelse(hospital$ANYOPER > 1,0,hospital$ANYOPER)
hospital$ANYOPER <- as.numeric(hospital$ANYOPER)

#Medicines prescribed at discharge - making assumption that 1 = yes and 2 = no
#what does -1,-7,-8,-9 mean? Different than blank?
#change na's to no?
hospital$DSCHPMED[hospital$DSCHPMED == 1] <- 1
hospital$DSCHPMED[hospital$DSCHPMED == 2] <- 0
hospital$DSCHPMED[is.na(hospital$DSCHPMED)] <- 0
hospital$DSCHPMED <- ifelse(hospital$DSCHPMED < 0,0,hospital$DSCHPMED)
hospital$DSCHPMED <- ifelse(hospital$DSCHPMED > 1,0,hospital$DSCHPMED)
hospital$DSCHPMED <- as.numeric(hospital$DSCHPMED)

#VA Facility flag
hospital$VAPLACE[hospital$VAPLACE == 1] <- 'Yes'
hospital$VAPLACE[hospital$VAPLACE == 0] <- 'No'
hospital$VAPLACE <- as.factor(hospital$VAPLACE)

hospital$year <- as.factor(hospital$year)

#cost amounts - mdtotal, mdpayment, fctotal, fcpayment, totalcharge, totalexpenditure - what do -1 records represent? change to $0?
hospital$mdpayment[hospital$mdpayment == -1] <- 0
hospital$mdtotal[hospital$mdtotal == -1] <- 0
hospital$fcpayment[hospital$fcpayment == -1] <- 0
hospital$fctotal[hospital$fctotal == -1] <- 0
hospital$totalcharge[hospital$totalcharge == -1] <- 0
hospital$totalexpenditure[hospital$totalexpenditure == -1] <- 0

str(hospital)
summary(hospital)
colnames(hospital)
nrow(hospital)
hospital_unique_records <- hospital %>% group_by(DUPERSID, year, EVNTIDX) %>% summarize(count=n())
nrow(hospital_unique_records)   # PK - DUPERSID, year, EVNTIDX
rm(hospital_unique_records)

# target variable creation
list <- c("480", "481", "482", "483", "484", "485", "486", "V09")
hospital_target <- rbind(hospital[hospital$icd1x %in% list,],
                         hospital[hospital$icd2x %in% list,],
                         hospital[hospital$icd3x %in% list,],
                         hospital[hospital$icd4x %in% list,])

hospital_target <- unique(hospital_target)
hospital_target$target <- 1
colnames(hospital_target)
hospital_target_eventid <- subset(hospital_target, select = c("DUPERSID","year","EVNTIDX","target"))
hospital_2 <- merge(hospital, hospital_target_eventid, by = c("DUPERSID","year","EVNTIDX"), all.x = TRUE)
hospital_2$target[hospital_2$target == 1] <- "Positive"
hospital_2$target[is.na(hospital_2$target)] <- "Negative"
hospital_2$target <- as.factor(hospital_2$target)
hospital_2$ERHEVIDX[!is.na(hospital_2$ERHEVIDX)] <- 1
hospital_2$ERHEVIDX[is.na(hospital_2$ERHEVIDX)] <- 0

hospital_2$num_nights[is.na(hospital_2$num_nights)] <- 0
colnames(hospital_2)

myvars <- c("DUPERSID", "EVNTIDX", "icd1x", "icd2x", "icd3x", "icd4x", "pro1x", "pro2x", "VAPLACE", "year", "mdtotal", "mdpayment", "fctotal", "fcpayment", "totalcharge", "totalexpenditure", "ERHEVIDX", "EMERROOM", "num_nights", "SPECCOND", "RSNINHOS", "ANYOPER", "DSCHPMED","target")

hospital_2 <- hospital_2[myvars]
summary(hospital_2)

rm(hospital,hospital_target,hospital_target_eventid)
```

## Creating a hospital table at patient - year level

```{r}
str(hospital_2)
hospital_3 <- hospital_2

hospital_3$VAPLACE <- as.character(hospital_3$VAPLACE)
hospital_3$VAPLACE[hospital_3$VAPLACE == 'Yes'] <- 1
hospital_3$VAPLACE[hospital_3$VAPLACE == 'No'] <- 0
hospital_3$VAPLACE <- as.numeric(hospital_3$VAPLACE)

hospital_3$target <- as.numeric(hospital_3$target)
hospital_3$target[hospital_3$target == 1] <- 0
hospital_3$target[hospital_3$target == 2] <- 1
hospital_3$target <- as.numeric(hospital_3$target)

hospital_3$patientid_year = paste(hospital_3$DUPERSID, hospital_3$year, sep='-')

# getting the data to patient-year level
hospital_4 <- hospital_3 %>% group_by(patientid_year) %>% summarize(visit_count=n(), 
                                                                    mdtotal = sum(mdtotal),
                                                                    fctotal = sum(fctotal),
                                                                    fcpayment = sum(fcpayment),
                                                                    totalcharge = sum(totalcharge),
                                                                    totalexpenditure = sum(totalexpenditure),
                                                                    target = max(target),
                                                                    VAFacility = max(VAPLACE),
                                                                    emer_related_stay = max(SPECCOND),
                                                                    emer_medication = max(DSCHPMED),
                                                                    operation = max(ANYOPER),
                                                                    num_nights = sum(num_nights),
                                                                    emer_room = max(EMERROOM),
                                                                    emer_rec = max(ERHEVIDX)
                                                                    )
table(hospital_4$target)

# patient procedures - 73 variables
proc1 = subset(hospital_3, select=c("patientid_year", "pro1x"))
colnames(proc1) <- c("patientid_year", "procedure")

proc2 = subset(hospital_3, select=c("patientid_year", "pro2x"))
colnames(proc2) <- c("patientid_year", "procedure")

proc <- rbind(proc1,proc2)
proc <- unique(proc)
rm(proc1,proc2)
proc <- subset(proc, proc$procedure != 'NA')

proc_list <- proc %>% group_by(procedure) %>% summarise(c = n())

proc_count <- proc %>% group_by(patientid_year) %>% summarize(proc_count=n()) 
proc$procedure <- paste("Proc_",proc$procedure, sep = '')
proc_pat <- dcast(proc,formula = patientid_year~procedure)

for (i in c(2:74)){
  proc_pat[i] <- ifelse(!is.na(proc_pat[i]),1,0) 
}

proc_pat <- merge(proc_count, proc_pat, by = "patientid_year")
proc_pat[is.na(proc_pat)] <- 0
patient_procedures <- proc_pat
hospital_4 <- merge(hospital_4, patient_procedures, by = "patientid_year", all.x = TRUE)

# patient emergency reasons - 6 variables
emer_reason_data = subset(hospital_3, select=c("patientid_year", "RSNINHOS"))
colnames(emer_reason_data) <- c("patientid_year", "emer_reason")
emer_reason_data$emer_reason <- paste("emer_reason_",emer_reason_data$emer_reason, sep = '')
emer_reason_data <- dcast(emer_reason_data,formula = patientid_year~emer_reason)

for (i in c(2:7)){
  emer_reason_data[i] <- ifelse(!is.na(emer_reason_data[i]),1,0) 
}
emer_reason_data[is.na(emer_reason_data)] <- 0
patient_emer_reasons <- emer_reason_data
hospital_4 <- merge(hospital_4, emer_reason_data, by = "patientid_year", all.x = TRUE)

table(hospital_4$target)
rm(proc_pat, proc, proc_count, proc_list, emer_reason_data, hospital_3)
```

## RX DATASET
```{r MEDICATIONS}
setwd(folder)
medications <- read.csv("rx0304_mod.csv", header = TRUE, sep=",",na.strings = c("-9", "-8","-7","-1"))

str(medications)
summary(medications)

colnames(medications) <- c(
"DUPERSID","RXRECIDX","LINKIDX","RXNAME","PharmType1","PharmType2",   "DIABFLG","RXICD1X","RXICD2X","RXICD3X","RXClass1", "RXClass2", "RXClass3","year",
"rxpayment", "rxexpenditure"
  )
medications$RXRECIDX <- as.character(medications$RXRECIDX)
medications$year <- as.factor(medications$year)

#pharmtype - need to find way to set all values not between 1 and 5 equal to NA then run codes below to change 1-5 values to Mail, Store, etc.
summary(medications$PharmType1)
medications$PharmType1 <- as.character(medications$PharmType1)
medications$PharmType1[medications$PharmType1 == 1] <- 'Mail'
medications$PharmType1[medications$PharmType1 == 2] <- 'Store'
medications$PharmType1[medications$PharmType1 == 3] <- 'Hospital'
medications$PharmType1[medications$PharmType1 == 4] <- 'Drugstore'
medications$PharmType1[medications$PharmType1 == 5] <- 'Online'
medications$PharmType1[is.na(medications$PharmType1)] <- 'Unknown'
medications$PharmType1 <- as.factor(medications$PharmType1)

#do same steps for pharmtype2 as pharmtype1
summary(medications$PharmType2)
medications$PharmType2 <- as.character(medications$PharmType2)
medications$PharmType2[medications$PharmType2 == 1] <- 'Mail'
medications$PharmType2[medications$PharmType2 == 2] <- 'Store'
medications$PharmType2[medications$PharmType2 == 3] <- 'Hospital'
medications$PharmType2[medications$PharmType2 == 4] <- 'Drugstore'
medications$PharmType2[medications$PharmType2 == 5] <- 'Online'
medications$PharmType2 <- as.factor(medications$PharmType2)

#diabflag - need to make sure this means that medicine is for diabetes
summary(medications$DIABFLG)
medications$DIABFLG <- as.character(medications$DIABFLG)
#medications$DIABFLG[medications$DIABFLG == '.'] <- 'No'
medications$DIABFLG[medications$DIABFLG == 0] <- 'No'
medications$DIABFLG[medications$DIABFLG == 1] <- 'Yes'
medications$DIABFLG <- as.factor(medications$DIABFLG)

#RXClass1,2,3 - same codes for each - change all values not in below list to NA
summary(medications$RXClass1)
medications$RXClass1 <- as.character(medications$RXClass1)
medications$RXClass1[medications$RXClass1 == 1] <- 'Antiinfective'
medications$RXClass1[medications$RXClass1 == 19] <- 'UnknownCategory'
medications$RXClass1[medications$RXClass1 == 20] <- 'Antineoplastics'
medications$RXClass1[medications$RXClass1 == 28] <- 'Biologicals'
medications$RXClass1[medications$RXClass1 == 40] <- 'Cardiovascular'
medications$RXClass1[medications$RXClass1 == 57] <- 'CNS'
medications$RXClass1[medications$RXClass1 == 81] <- 'Coagulation'
medications$RXClass1[medications$RXClass1 == 87] <- 'Gastrointestinal'
medications$RXClass1[medications$RXClass1 == 97] <- 'Hormones'
medications$RXClass1[medications$RXClass1 == 105] <- 'Misc'
medications$RXClass1[medications$RXClass1 == 115] <- 'Nutritional'
medications$RXClass1[medications$RXClass1 == 122] <- 'Respiratory'
medications$RXClass1[medications$RXClass1 == 133] <- 'Topical'
medications$RXClass1[medications$RXClass1 == 218] <- 'Alternative'
medications$RXClass1[medications$RXClass1 == 242] <- 'Psychotherapeutic'
medications$RXClass1[medications$RXClass1 == 254] <- 'Immonologic'
medications$RXClass1[medications$RXClass1 == 331] <- 'Radiologic'
medications$RXClass1[medications$RXClass1 == 358] <- 'Metabolic'
medications$RXClass1[is.na(medications$RXClass1)] <- 'Unknown'
medications$RXClass1 <- as.factor(medications$RXClass1)
summary(medications$RXClass1)

summary(medications$RXClass2)
medications$RXClass2 <- as.character(medications$RXClass2)
medications$RXClass2[medications$RXClass2 == 1] <- 'Antiinfective'
medications$RXClass2[medications$RXClass2 == 19] <- 'UnknownCategory'
medications$RXClass2[medications$RXClass2 == 20] <- 'Antineoplastics'
medications$RXClass2[medications$RXClass2 == 28] <- 'Biologicals'
medications$RXClass2[medications$RXClass2 == 40] <- 'Cardiovascular'
medications$RXClass2[medications$RXClass2 == 57] <- 'CNS'
medications$RXClass2[medications$RXClass2 == 81] <- 'Coagulation'
medications$RXClass2[medications$RXClass2 == 87] <- 'Gastrointestinal'
medications$RXClass2[medications$RXClass2 == 97] <- 'Hormones'
medications$RXClass2[medications$RXClass2 == 105] <- 'Misc'
medications$RXClass2[medications$RXClass2 == 115] <- 'Nutritional'
medications$RXClass2[medications$RXClass2 == 122] <- 'Respiratory'
medications$RXClass2[medications$RXClass2 == 133] <- 'Topical'
medications$RXClass2[medications$RXClass2 == 218] <- 'Alternative'
medications$RXClass2[medications$RXClass2 == 242] <- 'Psychotherapeutic'
medications$RXClass2[medications$RXClass2 == 254] <- 'Immonologic'
medications$RXClass2[medications$RXClass2 == 331] <- 'Radiologic'
medications$RXClass2[medications$RXClass2 == 358] <- 'Metabolic'
medications$RXClass2 <- as.factor(medications$RXClass2)
summary(medications$RXClass2)

summary(medications$RXClass3)
medications$RXClass3 <- as.character(medications$RXClass3)
medications$RXClass3[medications$RXClass3 == 1] <- 'Antiinfective'
medications$RXClass3[medications$RXClass3 == 19] <- 'UnknownCategory'
medications$RXClass3[medications$RXClass3 == 20] <- 'Antineoplastics'
medications$RXClass3[medications$RXClass3 == 28] <- 'Biologicals'
medications$RXClass3[medications$RXClass3 == 40] <- 'Cardiovascular'
medications$RXClass3[medications$RXClass3 == 57] <- 'CNS'
medications$RXClass3[medications$RXClass3 == 81] <- 'Coagulation'
medications$RXClass3[medications$RXClass3 == 87] <- 'Gastrointestinal'
medications$RXClass3[medications$RXClass3 == 97] <- 'Hormones'
medications$RXClass3[medications$RXClass3 == 105] <- 'Misc'
medications$RXClass3[medications$RXClass3 == 115] <- 'Nutritional'
medications$RXClass3[medications$RXClass3 == 122] <- 'Respiratory'
medications$RXClass3[medications$RXClass3 == 133] <- 'Topical'
medications$RXClass3[medications$RXClass3 == 218] <- 'Alternative'
medications$RXClass3[medications$RXClass3 == 242] <- 'Psychotherapeutic'
medications$RXClass3[medications$RXClass3 == 254] <- 'Immonologic'
medications$RXClass3[medications$RXClass3 == 331] <- 'Radiologic'
medications$RXClass3[medications$RXClass3 == 358] <- 'Metabolic'
medications$RXClass3 <- as.factor(medications$RXClass3)
summary(medications$RXClass3)/

medications_unique_records <- medications %>% group_by(DUPERSID, RXRECIDX, RXNAME, RXICD1X, rxpayment, year) %>% summarize(count=n())
nrow(medications_unique_records)   # PK - DUPERSID, RXRECID, year
rm(medications_unique_records)
summary(medications)

```

## MERGING DATASETS - hospital, demographics and conditions
```{r MERGING DATASETS}
hosp_demo <- merge(hospital_2, demographics_2, by = c("DUPERSID","year"))

conditions_4 <- conditions_3
conditions_4$patientid_year = paste(conditions_4$DUPERSID, conditions_4$year, sep='-')
conditions_4 = subset(conditions_4, select=c("patientid_year", "ICD9CODX"))
colnames(conditions_4) <- c("patientid_year", "condition")

conditions_5 <- subset(conditions_4, conditions_4$condition != 'NA')
#conditions_5$condition <- paste("cond_",conditions_5$condition, sep = '')
conditions_5 <- unique(conditions_5)
cond_count <- conditions_5 %>% group_by(patientid_year) %>% summarize(cond_count=n()) 
hospital_patientyr <- hospital_4 %>% group_by(patientid_year) %>% summarize(c=n()) 
patientyr <- as.data.frame(hospital_patientyr$patientid_year)
colnames(patientyr) <- c("patientid_year")
conditions_6 <- merge(patientyr, conditions_5, by = "patientid_year", all.x = TRUE)

icd_group <- read.csv("icd_list_final2.csv")
colnames(icd_group)
colnames(icd_group) <- c("condition","condition_desc")

conditions_6 <- merge(conditions_6,icd_group, by = "condition", all.x = TRUE )
conditions_6 <- subset(conditions_6, select = c("patientid_year", "condition_desc"))
colnames(conditions_6) <- c("patientid_year","condition")
summary(cond_count)
conditions_6$condition <- paste("cond_",conditions_6$condition, sep = '')
patient_condition <- dcast(conditions_6,formula = patientid_year~condition)

for (i in c(2:23)){
  patient_condition[i] <- ifelse(patient_condition[i] == 0,0,1) 
}

patient_conditions <- merge(patient_condition,cond_count, by = "patientid_year", all.x = TRUE)

rm(conditions_6, conditions_5, conditions_4, cond_count,patient_condition  )
```

## MERGING DATASETS - hospital, demographics, conditions and medications

# medications - PharmType1,2

```{r}
# PharmType1
summary(medications)
medications_1 <- medications
medications_12 <- medications

medications_1$patientid_year = paste(medications_1$DUPERSID, medications_1$year, sep='-')
medications_1 = subset(medications_1, select=c("patientid_year", "PharmType1"))
colnames(medications_1) <- c("patientid_year", "PharmType")

medications_12$patientid_year = paste(medications_12$DUPERSID, medications_12$year, sep='-')
medications_12 = subset(medications_12, select=c("patientid_year", "PharmType2"))
colnames(medications_12) <- c("patientid_year", "PharmType")

medications_2 <- rbind(medications_1,medications_12)
medications_2 <- unique(medications_2)
rm(medications_12, medications_1)

medications_2$PharmType <- paste(medications_2$PharmType, " : ",sep = '')
medications_2 <- subset(medications_2, medications_2$PharmType != 'NA : ')
medications_2$PharmType <- substr(medications_2$PharmType, 1, nchar(medications_2$PharmType)-3)
medications_2$PharmType <- paste("PharmType_",medications_2$PharmType, sep = '')

medications_3 <- dcast(medications_2,formula = patientid_year~PharmType)

for (i in c(2:7)){
  medications_3[i] <- ifelse(!is.na(medications_3[i]),1,0) 
}

patient_pharmatypes <- medications_3

rm(medications_3, medications_2)
```

# RXClass1,2,3

```{r}
# RXClass1
table(final_patientyr_data_2$target)
summary(medications)

medications_1 <- medications
medications_12 <- medications
medications_13 <- medications

medications_1$patientid_year = paste(medications_1$DUPERSID, medications_1$year, sep='-')
medications_1 = subset(medications_1, select=c("patientid_year", "RXClass1"))
colnames(medications_1) <- c("patientid_year", "RXClass")

medications_12$patientid_year = paste(medications_12$DUPERSID, medications_12$year, sep='-')
medications_12 = subset(medications_12, select=c("patientid_year", "RXClass2"))
colnames(medications_12) <- c("patientid_year", "RXClass")

medications_13$patientid_year = paste(medications_13$DUPERSID, medications_13$year, sep='-')
medications_13 = subset(medications_13, select=c("patientid_year", "RXClass3"))
colnames(medications_13) <- c("patientid_year", "RXClass")

medications_2 <- rbind(medications_1,medications_12,medications_13)
medications_2 <- unique(medications_2)
rm(medications_12, medications_1,medications_13)

medications_2$RXClass <- paste(medications_2$RXClass, " : ",sep = '')
medications_2 <- subset(medications_2, medications_2$RXClass != 'NA : ')
medications_2$RXClass <- substr(medications_2$RXClass, 1, nchar(medications_2$RXClass)-3)
medications_2$RXClass <- paste("RXClass_",medications_2$RXClass, sep = '')

medications_3 <- dcast(medications_2,formula = patientid_year~RXClass)

for (i in c(2:19)){
  medications_3[i] <- ifelse(!is.na(medications_3[i]),1,0) 
}

patient_class <- medications_3

rm(medications_3, medications_2)
```

## Medications 

```{r}
colnames(medications)
med1 <- medications
med1$patientid_year = paste(med1$DUPERSID, med1$year, sep='-')
med1 = subset(med1, select=c("patientid_year", "RXNAME"))
colnames(med1) <- c("patientid_year", "medication")

med2 <- subset(med1, med1$medication != 'NA')
med2$medication <- paste("med_",med2$medication, sep = '')
med2 <- unique(med2)
med_count <- med2 %>% group_by(patientid_year) %>% summarize(med_count=n()) 
med3 <- merge(patientyr, med2, by = "patientid_year", all.x = TRUE)

patient_med <- dcast(med3,formula = patientid_year~medication)

for (i in c(2:4217)){
  patient_med[i] <- ifelse(!is.na(patient_med[i]),1,0) 
}

patient_medications <- merge(med_count, patient_med, by = "patientid_year", all.y = TRUE)
rm(med1, med2, med3, patient_med, med_count  )
```



## Merging all datasets at patient-yr level


```{r} 
# hospital with demographics - hosp_demo_patientyr_level 

hospital_patientyr <- hospital_4
demographics_2$patientid_year = paste(demographics_2$DUPERSID, demographics_2$year, sep='-')
hosp_demo_patientyr_level <- merge(hospital_patientyr, demographics_2, by = "patientid_year")
table(hosp_demo_patientyr_level$target)
sum(hospital_patientyr$target)

# patient conditions with hosp_demo_patientyr_level - hosp_demo_cond_patientyr_level 
hosp_demo_cond_patientyr_level <- merge(hosp_demo_patientyr_level, patient_conditions, by = "patientid_year", all.x = TRUE)
hosp_demo_cond_patientyr_level[is.na(hosp_demo_cond_patientyr_level)] <- 0
table(hosp_demo_cond_patientyr_level$target)


# pharmacy type visited with hosp_demo_cond_med_patientyr_level - hosp_demo_cond_med_pt_patientyr 
hosp_demo_cond_pt_patientyr <- merge(patient_pharmatypes, hosp_demo_cond_patientyr_level, by = "patientid_year", all.y = TRUE)
hosp_demo_cond_pt_patientyr[is.na(hosp_demo_cond_pt_patientyr)] <- 0
table(hosp_demo_cond_pt_patientyr$target)

# Symptom Class with hosp_demo_cond_med_pt_patientyr - hosp_demo_cond_med_pt_cl_patienty
hosp_demo_cond_pt_cl_patientyr <- merge(patient_class, hosp_demo_cond_pt_patientyr, by = "patientid_year", all.y = TRUE)
hosp_demo_cond_pt_cl_patientyr[is.na(hosp_demo_cond_pt_cl_patientyr)] <- 0
table(hosp_demo_cond_pt_cl_patientyr$target)

final_patientyr_data_2 <- hosp_demo_cond_pt_cl_patientyr
final_patientyr_data_2[is.na(final_patientyr_data_2)] <- 0

folder="C:/Users/Lavanya/Google Drive/Group Project/Lavanya/FinalData"
setwd(folder)
write.csv(final_patientyr_data_2, "final_patientyr_data_2.csv", row.names = FALSE)

icd_df1 <- as.data.frame(hospital_2$icd1x)
colnames(icd_df1) <- c("icd")
icd_df2 <- as.data.frame(hospital_2$icd2x)
colnames(icd_df2) <- c("icd")
icd_df3 <- as.data.frame(hospital_2$icd3x)
colnames(icd_df3) <- c("icd")
icd_df4 <- as.data.frame(hospital_2$icd4x)
colnames(icd_df4) <- c("icd")
icd_df5 <- as.data.frame(conditions_3$ICD9CODX)
colnames(icd_df5) <- c("icd")
icd_df <- rbind(icd_df1,icd_df2,icd_df3,icd_df4,icd_df5)
rm(icd_df1,icd_df2,icd_df3,icd_df4,icd_df5) 
icd_df <- unique(icd_df)
folder="C:/Users/Lavanya/Google Drive/Group Project/Lavanya/FinalData"
setwd(folder)
write.csv(icd_df, "icd_list.csv", row.names = F)

```